---
## Hierarchical GLM model

Set some stan settings. 

```{r appaoa-stan_setup, eval=FALSE}
rstan::rstan_options(auto_write = TRUE,mc.cores = parallel::detectCores())
# ms <- read_csv("engdata.csv")
```

The model. Constraints:

* No negative slopes - we don't get worse at words. 
* Strong prior on slopes, they should have mean and sd around .25 (empirical)
* Much weaker prior on intercepts, though yoked their SDs (perhaps too tightly?


```{r appaoa-stan_model, eval=FALSE}
hierarchical_logit <-'
data {                            
  int<lower=1> W; // number of words
  int<lower=1> A; // number of ages
  vector[A] age; // subject ages
  int<lower=0> produces[W,A]; // count data
  int<lower=0> attempts[W,A]; // count data
}

parameters {
  real mu_i;             // intercept mean
  real<lower=0> sigma_i; // intercept SD
  
  real<lower=0> mu_s;             // slope mean
  real<lower=0> sigma_s; // slope SD

  vector[W] intercept; // word means
  vector[W] slope; // subject means
}

transformed parameters {
  matrix[W,A] p;

  for (w in 1:W) 
    for (a in 1:A)
      p[w,a] = intercept[w] + (slope[w] * age[a]);
}

model {
  mu_i ~ normal(0, 10); 
  sigma_i ~ normal(0, 1);

  mu_s ~ normal(0, .25);
  sigma_s ~ normal(0, .25);

  intercept ~ normal(mu_i, sigma_i);
  slope ~ normal(mu_s, sigma_s);
  
  for (w in 1:W) 
    for (a in 1:A)
      produces[w,a] ~ binomial_logit(attempts[w,a], p[w,a]);
}
'
```

Now reformat the data to stan format and compute. 

```{r appaoa-stan_prep, eval=FALSE}
model.data <- ms %>%
  ungroup %>%
  mutate(word = definition, 
         n = num_true,
         N = num_true + num_false) %>%
  # filter(word %in% c("daddy*","mommy*","no","bye")) %>%
  select(word, age, category, n, N)

ages <- unique(model.data$age)
n.words <- length(unique(model.data$word))

dat <- list(age = ages,
            produces = matrix(model.data$n, 
                              nrow=n.words, ncol=length(ages), byrow = TRUE),
            attempts = matrix(model.data$N, 
                              nrow=n.words, ncol=length(ages), byrow = TRUE),
            W = n.words, 
            A = length(ages))
```

```{r appaoa-stan_fit, eval=FALSE}
samps <- rstan::stan(model_code = hierarchical_logit, 
              cores = 4, 
              data = dat, iter = 200, warmup=100, chains = 4, 
              pars = c("mu_i", "sigma_i", "mu_s", "sigma_s", "slope", "intercept"))
save(samps, "data/appendix_aoa/aoa-samps.Rds")
```

Diagnostics.

```{r appaoa-stan-diagnostics, eval=FALSE}
load("data/appendix-aoa/aoa-samps.Rds")
rstan::traceplot(samps, pars = c("mu_i","sigma_i","mu_s","sigma_s"), 
                 inc_warmup = TRUE)
```

Explore parameters. 

```{r appaoa-stan_params, eval=FALSE}
coefs <- data.frame(summary(samps)$summary)
coefs$name <- rownames(coefs)

word_ids <- model.data %>% 
  group_by(word) %>% 
  summarise(category = category[1]) %>%
  mutate(word_id = 1:n())

words <- coefs %>% 
  filter(str_detect(name, "slope") | str_detect(name, "intercept")) %>%
  separate(name, c("variable", "word_id"), "\\[") %>%
  mutate(word_id = as.numeric(str_replace(word_id, "]", ""))) %>%
  left_join(word_ids) %>%
  select(mean, variable, word, category) %>%
  spread(variable, mean) %>%
  mutate(aoa = intercept / slope) %>%
  arrange(aoa) %>%
  mutate(word = factor(word, 
                        levels = word,
                        labels = word))
```

and plot again:

```{r appaoa-stan_plot, eval=FALSE}
age_range <- 0:36
preds <- words %>%
  group_by(word, category) %>%
  do(data.frame(age = age_range, 
                pred.prop = boot::inv.logit(.$intercept + age_range * .$slope))) %>%
  left_join(model.data) %>%
  mutate(prop = n/N) 

ggplot(filter(preds, word %in% target_words), aes(x = age, y = prop)) + 
  geom_point() + 
  facet_wrap(~word) + 
  geom_line(aes(x = age, y = pred.prop))
```

Check the histogram.

```{r appaoa-stan_hist, eval=FALSE}
hglm.aoas <- words %>%
  mutate(aoa = -intercept / slope)

qplot(aoa, geom = "blank", data = hglm.aoas) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density(col = "red")
```

What are the crazy ones? 

```{r appaoa-stan_outliers, eval=FALSE}
hglm.aoas$word[hglm.aoas$aoa <10]
```

Don't know what's going on with these but everything else looks good. 
