# Morphological (Over-)Generalization {#overregularization}

```{r ovreg-prep_data, eval=FALSE}
grammar_data_coded <- read_feather("data/grammar/grammar_data_coded.feather")

overreg_data <- grammar_data_coded %>%
  filter(type %in% c("word_endings_nouns", "word_endings_verbs")) %>%
  mutate(value = if_else(is.na(value), FALSE, value),
         type = fct_recode(type, "nouns" = "word_endings_nouns",
                           "verbs" = "word_endings_verbs"))

write_feather(overreg_data, "data/overreg_data.feather")
```

```{r ovreg-load_data}
overreg_data <- read_feather("data/grammar/overreg_data.feather")

overreg_longs <- overreg_data %>%
  left_join(admins) %>%
  filter(longitudinal)
```


## Item trajectories

First, look at the trajectories of how many children are reported to produce each item.

```{r ovreg-byitem}
overreg_item_summary <- overreg_data %>%
  group_by(language, age, type, definition) %>%
  summarise(num_true = sum(value, na.rm = TRUE),
            total = n(),
            num_false = total - num_true,
            prop = num_true / total)

plt_item_traj <- function(lang, typ) {
  overreg_item_summary %>%
    filter(language == lang, type == typ) %>%
    ggplot(aes(x = age, y = prop)) +
      facet_wrap(~definition, ncol = 5) +
      geom_smooth(method = "loess") +
      labs(x = "Age (Months)", y = "Proportion of children producing")
}
```

__English nouns__
```{r ovreg-plt_traj_eng_nouns, dependson="ovreg_byitem", fig.height=5}
plt_item_traj("English (American)", "nouns")
```

__English verbs__
```{r ovreg-plt_traj_eng_verbs, dependson="ovreg_byitem", fig.height=11}
plt_item_traj("English (American)", "verbs")
```

__Norwegian nouns__
```{r ovreg-plt_traj_nor_nouns, dependson="ovreg_byitem", fig.height=5}
plt_item_traj("Norwegian", "nouns")
```

__Norwegian verbs__
```{r ovreg-plt_traj_nor_verbs, dependson="ovreg_byitem", fig.height=11}
plt_item_traj("Norwegian", "verbs")
```

Many of these looks flat over development, some go up, some have U-shape or other curves. Since these are averaging over many children, they might be obscuring the developmental trajectories of individual children.

## Items by trajectory types

Next, take only the longitudinal data and categorize each child's trajectory for each item by whether over age, they say it never, always, start and then continue, start and then stop, stop and then restart, stop, or something else. For each item, look at the proportions of children showing each of these patterns.

```{r ovreg-traj_props, eval=FALSE}
overreg_long_kids <- overreg_longs %>%
  select(language, type, definition, sex, original_id, age, value) %>%
  group_by(language, type, definition, sex, original_id) %>%
  filter(n() > 1) %>%
  arrange(age) %>%
  nest()

overreg_long_kids_coded <- overreg_long_kids %>%
  mutate(coded = map(data, function(kid_data) {
    vals <- kid_data$value[c(TRUE, diff(kid_data$value) != 0)]
    data_frame(
      always_says = length(vals) == 1 && vals == TRUE,
      never_says = length(vals) == 1 && vals == FALSE,
      starts_saying = length(vals) == 2 && vals == c(FALSE, TRUE),
      stops_saying = length(vals) == 2 && vals == c(TRUE, FALSE),
      starts_stops_saying = length(vals) == 3 && vals == c(FALSE, TRUE, FALSE),
      stops_starts_saying = length(vals) == 3 && vals == c(TRUE, FALSE, TRUE),
      other = !any(always_says, never_says, starts_saying, stops_saying,
                   starts_stops_saying, stops_starts_saying)
    )
  })) %>%
  select(-data) %>%
  unnest()

# overreg_long_kids_coded %>%
#   gather(measure, value, never_says, starts_saying, always_says, stops_saying,
#          starts_stops_saying, stops_starts_saying, other) %>%
#   group_by(language, definition, original_id) %>%
#   summarise(n_true = sum(value)) %>%
#   filter(n_true != 1)

overreg_long_words <- overreg_long_kids_coded %>%
  group_by(language, type, definition) %>% #, sex) %>%
  summarise(total = n(),
            always = sum(always_says) / total,
            never = sum(never_says) / total,
            starts = sum(starts_saying) / total,
            stops = sum(stops_saying) / total,
            starts_stops = sum(starts_stops_saying) / total,
            stops_starts = sum(stops_starts_saying) / total,
            other = sum(other) / total) %>%
  gather(measure, value, never, starts, always, stops, starts_stops,
         stops_starts, other) %>%
  group_by(language, type, measure) %>%
  arrange(desc(value)) %>%
  ungroup() %>%
  mutate(measure = fct_inorder(measure),
         definition = fct_inorder(definition))

write_feather(overreg_long_words, "data/grammar/overreg_long_words.feather")
```

```{r ovreg-plt_props}
overreg_long_words <- read_feather("data/grammar/overreg_long_words.feather")

plot_type_props <- function(typ) {
  overreg_long_words %>% filter(type == typ) %>%
    ggplot(aes(x = definition, y = value, fill = measure)) +
      facet_grid(language ~ ., scales = "free") +
      coord_flip() +
      geom_col(position = "stack") +
      scale_fill_ptol(name = "") +
      scale_y_continuous(expand = c(0, 0)) +
      labs(x = "", y = "Proportion of children") +
      theme(legend.position = "top") +
      guides(fill = guide_legend(nrow = 1, reverse = TRUE))
}
```

__Nouns__
```{r ovreg-plt_props_nouns, dependson="ovreg_plt_props", fig.height=5.5}
plot_type_props("nouns")
```

__Verbs__
```{r ovreg-plt_props_verbs, dependson="ovreg_plt_props", fig.height=11}
plot_type_props("verbs")
```


## Individual children

The trajectory types above give a course idea of developmental change. Next, let's look at individual children rather than individual items, and characterize each child's developmental progression.

There are too many children overall to actually look at thei individual curves, so let's find ways of categorizing them. In English, there are only two ages in the longitudinal data, so we can each trajectory value (as in number of items produced at each age) and count how many children fall into each one.

```{r ovreg-bykid, eval=FALSE}
overreg_kid_summary <- overreg_longs %>%
  group_by(language, sex, original_id, type, definition) %>%
  filter(n() > 1) %>%
  arrange(age) %>%
  group_by(language, age, sex, original_id, type) %>%
  summarise(num_produces = sum(value),
            num_total = n(),
            num_no = num_total - num_produces)

overreg_traj <- overreg_kid_summary %>%
  group_by(language, sex, type, original_id) %>%
  nest() %>%
  mutate(traj = map_chr(data, function(d) {
    paste(paste(d$age, collapse = ","), paste(d$num_produces, collapse = ","),
          paste(d$num_total, collapse = ","), sep = ";")
  })) %>%
  count(language, sex, type, traj) %>%
  mutate(data = map(traj, function(dstr) {
    dparts <- str_split(dstr, ";")[[1]]
    data_frame(age = as.numeric(unlist(str_split(dparts[1], ","))),
               num_produces = as.numeric(unlist(str_split(dparts[2], ","))),
               num_total = as.numeric(unlist(str_split(dparts[3], ","))))
  })) %>%
  unnest() %>%
  mutate(prop = num_produces / num_total)

overreg_traj_bin <- overreg_traj %>%
  filter(n > 1) %>%
  mutate(n_bin_str = n %>%
           cut(breaks = c(1, 10, 100, 1000), labels = c(1, 10, 100)) %>%
           as.character(),
         n_bin = n_bin_str %>% as.numeric(),
         n_bin_str = map_chr(n_bin, ~paste(.x, .x * 10, sep = " – ")))

write_feather(overreg_traj_bin, "data/grammar/overreg_traj_bin.feather")
```

```{r ovreg-plt_bins_eng, fig.height=6}
overreg_traj_bin <- read_feather("data/grammar/overreg_traj_bin.feather")

ggplot(filter(overreg_traj_bin, language == "English (American)"),
       aes(x = age, y = prop, colour = n_bin_str)) +
  facet_grid(type ~ sex) +
  geom_line(aes(group = traj, size = n)) +
  geom_point() +
  scale_colour_ptol(name = "Number of children") +
  scale_size(guide = FALSE, range = c(0.2, 3)) +
  scale_x_continuous(name = "Age (months)", breaks = c(16, 28)) +
  scale_y_continuous(name = "Proportion of overregularization items") +
  theme(legend.position = "top")
```

```{r ovreg-models, eval=FALSE}
kid_model <- function(kid_data) {
  glm(cbind(num_produces, num_no) ~ age + I(age ^ 2), data = kid_data,
      family = "binomial")
}

overreg_models <- overreg_kid_summary %>%
  filter(language == "Norwegian") %>%
  group_by(language, sex, original_id, type) %>%
  nest() %>%
  mutate(model = map(data, kid_model),
         coefs = map(model, tidy))

overreg_coefs <- overreg_models %>%
  mutate(switch = map_lgl(coefs, function(coef) {
    "I(age^2)" %in% coef$term &&
    filter(coef, term == "I(age^2)")$p.value < 0.05 &&
      filter(coef, term == "I(age^2)")$estimate < 0
  })) %>%
  select(-data, -model) %>%
  unnest()

overreg_predictions <- overreg_models %>%
  mutate(augmented = map(model, function(m) {
    m_aug <- m %>% augment(type.predict = "response")
    m_y <- m_aug$cbind.num_produces..num_no. %>% as_data_frame()
    m_aug %>%
      select(-cbind.num_produces..num_no.) %>%
      bind_cols(m_y)
  }))

overreg_fits <- overreg_predictions %>%
  select(-data, -model, -coefs) %>%
  unnest()

overreg_joined <- overreg_kid_summary %>%
  filter(language == "Norwegian") %>%
  left_join(overreg_coefs) %>%
  select(-term, -estimate, -std.error, -statistic, -p.value) %>%
  distinct() %>%
  left_join(overreg_fits) %>%
  mutate(prop = num_produces / num_total)

write_feather(overreg_joined, "data/grammar/overreg_joined.feather")
```

```{r ovreg-joined}
overreg_joined <- read_feather("data/grammar/overreg_joined.feather")

overreg_switch <- filter(overreg_joined, switch)

min_age_n <- 8
dense_kids <- overreg_joined %>%
  group_by(language, sex, type, original_id) %>%
  filter(n() >= min_age_n)
```

In Norwegian, there are many more ages measured, so this sort of unique trajectory value analysis is impossible.

First, what are the trajectories of the most densely sampled Norwegian children, i.e. ones with CDI administrations at least `r min_age_n` different ages?

```{r ovreg-plt_dense_emp, dependson="ovreg_joined", fig.height=6}
ggplot(dense_kids, aes(x = age, y = prop, colour = original_id)) +
  facet_grid(type ~ sex) +
  geom_line(aes(group = original_id)) +
  scale_colour_ptol(guide = FALSE) +
  scale_x_continuous(name = "Age (months)") +
  scale_y_continuous(name = "Proportion of overregularization items")
```

Some of these are fairly clearly a U-shaped trajectory, i.e. they have a period of increasing overregularization followed by a period of decreasing overregularization. The discover which children show a U-shape, let's fit a logistic regression model for each child predicting whether they produce each word from a linear and quadratic effect of their age. U-shaped children are taken to be those for whom the quadratic term is negative and statistically significant.

__U-shaped children's empirical proportions__
```{r ovreg-plt_switch_emp, dependson="ovreg_joined", fig.height=6}
ggplot(overreg_switch, aes(x = age, y = prop, colour = original_id)) +
  facet_grid(type ~ sex) +
  geom_line(aes(group = original_id)) +
  scale_colour_pander(guide = FALSE) +
  scale_x_continuous(name = "Age (months)") +
  scale_y_continuous(name = "Proportion of overregularization items",
                     limits = c(0, 1))
```

__U-shaped children's model fits__
```{r ovreg-plt_switch_fit, dependson="ovreg_joined", fig.height=6}
ggplot(overreg_switch, aes(x = age, y = .fitted, colour = original_id)) +
  facet_grid(type ~ sex) +
  geom_line(aes(group = original_id)) +
  scale_colour_pander(guide = FALSE) +
  scale_x_continuous(name = "Age (months)") +
  scale_y_continuous(name = "Proportion of overregularization items",
                     limits = c(0, 1))
```
