# Morphology, Grammar, and the Lexicon^[Material in this chapter first reported in @braginsky2015.] {#grammar}

How does abstract structure emerge during language learning? On some accounts, children's early syntax emerges from direct generalizations from particular lexical items, while on others, syntactic structure is acquired independently and follows its own timetable. CDI data can help us decide between these two views. In this chapter, we summarize the state of grammatical development across languages (noting the challenges posed by radically different representations of grammar across CDI forms). We also replicate and generalize analyses linking grammatical generalization to children’s vocabulary size. We end by investigating the idea that that age modulates the relationship between grammar and the lexicon.

```{r gram-prep_data, child="_grammar_lex.Rmd", eval=FALSE}
```

```{r gram-load_data}
grammar_summary <- read_feather("data/grammar/grammar_summary.feather")
grammar_data_coded <- read_feather("data/grammar/grammar_data_coded.feather")

grammar_combine <- grammar_summary %>%
  filter(coded_type == "combine")

grammar_morphosyntax <- grammar_summary %>%
  ungroup() %>%
  filter(coded_type %in% c("complexity", "word_forms")) %>%
  mutate(coded_type = factor(coded_type,
                             levels = c("word_forms", "complexity"),
                             labels = c("Word Forms", "Complexity"))) %>%
  filter(language != "French (France)")  # TODO!!!

num_langs <- n_distinct(grammar_summary$language)
```


## Introduction

For many children, their first words are spoken in isolation. While these single word utterances sometimes seem to be picking out objects in the world (e.g., *ball!*), others seem to convey more complex ideas or desires (e.g., *up!* for *Mommy, pick me up!*). But by two years of age, many children have acquired a large repertoire of words, and are beginning to use them in two- or three-word combinations (e.g., *Mommy up!* or *kitty sleep here*). These utterances will gradually increase in length and complexity in various ways, forming sentences that increasingly reflect the grammatical structure of their native language (e.g., *Mommy, the kitty is sleeping here*). Children also begin to add more verbs, adjectives and other predicates to their working vocabularies (see Chapter \@ref(categories)), and substantively increase their use of prepositions, articles and other closed class forms that do grammatical work, including the productive use of inflectional morphemes (e.g., English past tense *ed* or *ing*).

Understanding the origins of grammar is critical because children's ability to use morphosyntactically-rich language is thought to reflect the uniquely-human mental machinary that enables speakers to produce novel utterances that have never been heard in the input [@berko1958; @pinker1991]. The questions surrounding the development of grammar are challenging. How do abstract morphosyntactic structures emerge during language learning? What mechanisms underlie the formation of generalizations that support such inferences and allow children to apply them during language production? Does an understanding of the abstract rule-structure of language emerge from the interactions of individual words, or is that structure independently acquired and represented separately?

Broadly speaking, theoretical views on grammatical development generally take one of two forms. On nativist theories like Principles and Parameters [@chomsky1981;@baker2005], grammar emerges independently from lexical knowledge following its own, largely maturational, timetable.  Moreover, grammatical regularities are mentally represented in a format that is distinct from that used by the lexical system. In contrast, according to lexicalist theories, mental representations of morphosyntactic structure generally emerge from graded generalizations on the basis of lexical items, and at least early in development, there may be little or no representation of morphosyntactic rules or regularities *per se* [@tomasello2003, @elman1996]. Even when syntactic structures are eventually represented, these representations are directly related to more concrete lexical structures [@bannard2009].

Historically, the study of individual differences has been critial to this debate.  While variation in word learning is generally uncontroversial, individual differences in grammatical development are less clearly predicted under a universalist, nativist perspective.  In contrast, lexicalist theories predict that variation in grammatical development should be tightly yoked to variation in lexical development [@bates1999]. Research has shown that, as with lexical development, there is sizeable variation in exactly when and how children move into using more grammatically complex utterances in their everyday speech.  While some children use primarily multi-word phrases and many closed class forms by 24 months, other children are still primarily producing nouns in single word utterances at that same age [e.g., @bates1988; @bates1999].  Moreover, there is also variation in the kinds of multi-word utterances that children produce.  For example, some children build up sentences from individual words (e.g., *want dat!*), whereas other children seem to produce utterances that reflect "unanalyzed" chunks of more complex speech (e.g., *iwantdodat!*).

Associations between individual differences in lexical and grammatical development have been robustly substantiated in the literature. In the original norming data from the English CDI: Words & Sentences, children with more sophisticated grammatical productions were also those children with the largest vocabularies [@bates1994]. Using that same dataset, @marchman1994 found that size of verb vocabulary was concurrently related to children's overregularization of past tense inflections (e.g., *daddy goed*), productions that are viewed as a major milestone in the development of grammatical rule-based knowledge. Links between lexical development and grammar have also been reported longitudinally [@bates1988; @bates1997], in late talkers [e.g., @paul1996;@rescorla2000;@rescorla1997;@thal1997], early talkers [@thal1996;@thal1997], and children with neurodevelopmental disorders, such as Williams syndrome [e.g., @singerharris1997]. Similar relationships have also been demonstrated in many other languages, including Slovenian [@marjanovic-umek2013], Hebrew [@maital2000], Icelandic [@thordardottir2002], Italian [@caselli1999;@devescovi2005], Bulgarian [@andonova2015], Finnish [@stolt2009], Spanish [@mariscal2012; @thal2000], and German [@szagun2006]. 

Finally, and perhaps most intriguingly, in behavioral genetic studies of monozygotic and dizygotic twins, the relation between lexical and grammatical level has been found to be strongly heritable [@dale2000; @dionne2003]. In other words, even though genetic factors contribute relatively weakly to each aspect of language as assessed individually, the genetic factors that influence lexical growth are the same as those that influence grammatical growth, perhaps operating in a bidirectional manner.

While these studies substantiate that vocabulary and grammar development are strongly associated developmentally, the interpretation of these relations is still under debate. Some researchers have interpreted these links to suggest that domain-general learning mechanisms guide the child’s construction of a working linguistic system at many different levels, in this case, learning words and learning grammatical rules [e.g., @elman1996].  As @bates1987 proposed many years ago, “the native speaker learns to map phrasal configurations onto propositions, using the same learning principles and representational mechanisms needed to map single words onto their meanings” (p. 163). Other proposals suggest that the process of learning words involves learning both their lexical-semantic and their morphosyntactic properties (e.g., in what constructions they can legally appear and what inflectional morphemes are required), and that grammatical knowledge is generally built up on a case-by-case basis [@tomasello2003]. Early word combinations are often highly routinized and situation specific, suggesting that learning grammar, like word learning, is guided by learning mechanisms that are item specific and frequency dependent. It is only later that grammatical structures become encoded in terms of their abstract syntactic form [e.g., @lieven1997; @tomasello2003]. Yet other accounts view the relation as reflecting mechanisms that operate in the opposite direction. On these views, grammatical analysis is a driving force behind word learning, such that the process of analyzing sentences into their constituent grammatical parts facilitates the further acquisition of lexical-semantic knowledge [@anisfeld1998;@naigles1990].  Finally, other studies have proposed that the lexical-grammar relations may not be as direct as previously proposed, actually being driven by common third-party influences such as the speech that children hear [@hoff2017].

## Goals of the current analyses

In this chapter, we explore relations between estimates of children's vocabulary size based on the vocabulary checklist and responses on other sections of the Words and Sentences instruments. Many versions of the instruments provide indices of grammar learning by asking about children's use of inflected forms (e.g., *walked*), children's use of overgeneralizations (e.g., *goed*), and the complexity of their multi-word combinations (e.g., *kitty sleeping* / *kitty is sleeping*).  While many studies have examined associations between lexical and grammatical development crosslinguistically, the scope and power of these early studies were limited, with few opportunities for direct comparisons of the nature or extent of these relations across multiple languages at the same time. In contrast, our data allow analyses of lexical-grammar relations with enhanced statistical power and broader cross-linguistic representation.  

In addition, we explore a hypothesis that was not explicitly tested in these earlier studies: that there remains age-related variance in grammatical development unexplained by vocabulary development. While the overall relationship between grammar and the lexicon provides support for lexicalist theories, the identification of age-related variance would suggest the presence of developmental processes that regulate grammar learning, above and beyond those captured by measures of vocabulary size. Such age-related processes could be either maturational or experiential, and either domain-general (like working memory) or language-specific (like grammatical competency). Importantly, since both nativist and constructivist theories could in principle predict age-linked variance in grammatical development, our goal is not to differentiate these theories, but instead to test this novel prediction and explore its implications for future work on understanding the processes of grammatical development.

An additional contribution of work is that, due to the size of our dataset, we are able to make more fine-grained distinctions than the initial cut between grammar and the lexicon. In particular, we distinguish morphology from multi-word syntax, since morphological generalizations might be more specifically dependent on vocabulary size than those requiring more global, sentence-level syntactic regularities. <!--Similarly, we distinguish age-related contributions to different parts of the vocabulary. Lexical items like verbs and even more so function words require some syntactic information to learn [@gleitman1990] and hence might be more linked to age-related factors that extend beyond vocabulary size.-->


## Methods

In all `r num_langs` languages included in these analyses, the CDI forms contain both vocabulary checklists and other questions relevant to the child's linguistic development. All of the data reported here come from Words & Sentences type forms, administered to children ages `r min(grammar_data_coded$age)`-- `r max(grammar_data_coded$age)` months (most in the 16--30 month range). In addition to the vocabulary checklist items, these forms typically contain a single item asking whether the child is combining words yet at all; Word Form section, which asks whether the child produces each of around 30 morphologically inflected forms of nouns and verbs (e.g., _feet_, _ran_); and a Complexity section, which asks whether the child's speech is most similar to the syntactically simpler or more complex versions of around 40 sentences (e.g., _two foot / two feet_, _there a kitty / there's a kitty_). 

Importantly, each instrument for languages other than English is not just a translation of the English form, but rather was constructed and normed to reflect the lexicon and grammar of that language. Thus, there are substantial differences in the content of these items and their coverage of different morphological and grammatical phenomena. The major commonality is that the form developers believed that they provided a good survey of important developmental phenomena in their language. 

Word Form items can be browsed in the table below:

```{r gram-item_tables}
grammar_items <- grammar_data_coded %>%
  ungroup() %>%
  distinct(language, form, coded_type, definition)
  
grammar_items %>%
  filter(coded_type == "word_forms") %>%
  datatable(rownames = FALSE, filter = "bottom", style = "bootstrap",
            options = list(pageLenth = 20))
```

Complexity items are shown in the table below:

```{r gram-complexity_table}
grammar_items %>%
  filter(coded_type == "complexity") %>%
  datatable(rownames = FALSE, filter = "bottom", style = "bootstrap",
            options = list(pageLenth = 20))
```

To analyze lexical and grammatical development, we derive several measures. Each child's Vocabulary Size is computed as the proportion of words on the corresponding CDI form that the child is reported to produce. Similarly, each child's Word Form score is the proportion of word forms they are reported to produce, and their Complexity score the proportion of complexity items for which they are reported to use the more complex form. We compute all of these quantities as proportions to make the scales comparable across languages.

## Results

We present four sets of results. First, we show analyses of the "combines" item, which is a binary item in which parents indicate whether their child is combining words. Second, we give analyses of the relationship between vocabulary size and Word Form/Complexity items. Third, we follow up on a pattern found in the "combines" item, namely age-related modulation of the grammar-lexicon relationship. Finally, we investigate the degree to which the age-related pattern is found in individual items. 

### Combine

```{r gram-combine_plot}
combine_compare <- grammar_combine %>%
  gather(predictor, value, age, production) %>%
  mutate(predictor = factor(
    predictor, levels = c("age", "production"),
    labels = c("Age (months)", "Productive vocabulary size")))

# slovak combine item is "your child uses 'fancy words' like bubble blower, brush and comb' - so I am excluding it. -MCF

ggplot(filter(combine_compare, language != "Slovak"),
       aes(x = value, y = positive, colour = language)) +
  facet_wrap(~predictor, scales = "free_x") +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              se = FALSE) +
  scale_colour_solarized(name = "") +
  labs(x = "", y = "Combines Words (Probability)") +
  theme(legend.position = "bottom")
```
The plot above shows the probability of a parent checking that their child combines words, plotted by the child's chronological age (left) and raw productive vocabulary size (right). As can be seen, across `r length(unique(combine_compare$language))-1 #slovak` languages, there is some consistency in the chronological trajectories for this item. By 24 months, around 75% of children are reported to be combining words, though this estimate is substantially earlier in Quebec French. One possibility is that the phrasing of the "combines" item contributes -- some forms (including Quebec French, but also Norwegian and Danish) give examples of simple combinations, which could encourage earlier reporting.  

Vocabulary-related trajectories were more variable, however. In general, children who were marked as combining had vocabularies larger than around 100 words. However, as noted in Chapter \@ref(vocab), raw Mandarin vocabulary in the WS form is unusually high, but the "combines" item does not appear to be comparably accelerated. Thus, Mandarin children appear to be producing words only after producing substantially more vocabulary items. On the opposite side, children learning Quebec French and Korean were reported to be combining with quite small vocabularies. 

To investigate the quantitative relationship between word combination (as measured with this item), age, and vocabulary, we fit a linear mixed effects model predicting combination as a function of vocabulary (proportion), age, and their interaction. We also included random intercepts by language and random slopes for both vocabulary and age. Coefficient estimates are shown below:

```{r gram-combines_lmer, eval=FALSE}
mod <- glmer(positive ~ production_prop * age +
               (production_prop + age | language), 
             family = "binomial", 
             data = filter(grammar_combine,  language != "Slovak"))
#full interaction doesn't converge.
save(file = "data/grammar/grammar_combines_model.Rds", mod)
```

```{r gram-coef_table}
load("data/grammar/grammar_combines_model.Rds")

knitr::kable(summary(mod)$coefficients, digits = 3)
```


This model shows an extremely large effect of vocabulary, with a relatively smaller amount of variance due to age. In addition, there was a substantial negative interaction of vocabulary and age, reflecting that older children were more likely to be combining words, even with less vocabulary. This result parallels others reported below suggesting that there are age-related components in grammatical performance that are unaccounted for by vocabulary. All coefficients were highly significant due to the large amount of data. 

Overall, although there was some cross-linguistic variation -- perhaps due to true variation and perhaps due to idiosyncrasies of individual forms or datasets -- word combination emerged around 24 months and 100 words for most children. 

### Grammar and lexicon relationship

```{r gram-vocab_models}
vocab_model <- function(inst_data) {
  lm(prop ~ I(production_prop ^ 3) + I(production_prop ^ 2) +
       production_prop + 0, data = inst_data)
}

vocab_models <- grammar_morphosyntax %>%
  group_by(language, instrument, coded_type) %>%
  nest() %>%
  mutate(
    model = map(data, vocab_model),
    rsq = map_dbl(model, ~summary(.x)$r.squared),
    rsq_print = sprintf("r² = %.2f", rsq)
  )

cvs <- vocab_models %>%
  group_by(coded_type) %>%
  summarise(cv = sd(rsq) / mean(rsq)) %>%
  mutate(foo = "and Productive Vocabulary (r²)") %>%
  unite(measure, coded_type, foo, sep = " ")

write_feather(cvs, "data/cvs/grammar_lex.feather")
```

```{r gram-grammar_data_plot}
plot_vocab_type <- function(type) {
  ggplot(filter(grammar_morphosyntax, coded_type == type),
         aes(x = production_prop, y = prop)) +
    geom_jitter(alpha = 0.2, size = 0.75, shape = 1) +
    facet_wrap(~language) +
    geom_smooth(method = "lm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x + 0,
                size = 1, se = FALSE) +
    geom_text(aes(label = rsq_print), x = 0.15, y = 1, size = 3, family = .font,
              data = filter(vocab_models, coded_type == type)) +
    labs(x = "Productive vocabulary size (proportion of items)",
         y = "Grammar score (proportion of items)",
         title = type)
}
```

We next examine the correlation between the proportion of Word Form/Complexity items completed and the proportion of vocabulary items completed. First reported by @bates1994, these correlations are extremely robust, and can be observed in essentially all of our datasets. Our first plot shows this relation for Word Form items. We fit linear regressions predicting vocabulary as a function of linear, quadratic, and cubic predictors (subtracting the intercept to ensure that the function passed through the origin). The total $r^2$ for these relationships ranged from `r min(filter(vocab_models, coded_type == "Word Forms")$rsq)` to `r max(filter(vocab_models, coded_type == "Word Forms")$rsq)`.


```{r gram-grammar_plot_wordform}
plot_vocab_type("Word Forms")
```

Complexity items show the same relationship, typically with equal or greater strength (depending on data density and number of items). $r^2$ values varied from `r min(filter(vocab_models, coded_type == "Complexity")$rsq)` to `r max(filter(vocab_models, coded_type == "Complexity")$rsq)`.


```{r gram-grammar_plot_complexity}
plot_vocab_type("Complexity")
```

Overall, these data provide strong cross-linguistic support to the contention of @bates1994 and others that the emergence of grammatical competence in production is related across individuals to the size of the productive vocabulary. 

### Age effects

```{r gram-grammar_age_models}
vocab_step <- 0.01

inst_model <- function(inst_data) {
  lm(prop ~ I(production_prop ^ 3) * age + I(production_prop ^ 2) * age +
       production_prop * age + 0, data = inst_data)
}

inst_predictions <- function(inst_data, inst_model) {
  data_shape <- cross_df(list(age = seq(min(inst_data$age), max(inst_data$age)),
                              production_prop = seq(0, 1, vocab_step)))
  augment(inst_model, newdata = data_shape, type.predict = "response")
}

grammar_models <- grammar_morphosyntax %>%
  group_by(language, instrument, coded_type) %>%
  nest() %>%
  mutate(
    model = map(data, inst_model),
    rsq = map_dbl(model, ~summary(.x)$r.squared),
    predictions = map2(data, model, inst_predictions)
  )

num_age_admins <- grammar_morphosyntax %>%
  ungroup() %>%
  distinct(language, instrument, age, data_id) %>%
  count(instrument, age)

grammar_fits <- grammar_models %>%
  select(language, instrument, coded_type, predictions) %>%
  unnest() %>%
  group_by(language, instrument) %>%
  mutate(age_group = cut(age, 2, labels = c("younger", "older"))) %>%
  left_join(num_age_admins) %>%
  group_by(language, instrument, coded_type, age_group, production_prop) %>%
  summarise(mean_fitted = weighted.mean(.fitted, n))

grammar_morphosyntax <- grammar_morphosyntax %>%
  group_by(instrument) %>%
  mutate(age_group = cut(age, 2, labels = c("younger", "older")))
```

```{r gram-grammar_age_plots}
grammar_age_plot <- function(type) {
  ggplot(filter(grammar_morphosyntax, coded_type == type),
         aes(x = production_prop, y = prop, colour = age_group)) +
    facet_wrap(~language) +
    geom_jitter(alpha = 0.1, size = 0.70, shape = 1) +
    geom_line(aes(y = mean_fitted), size = 0.9,
              data = filter(grammar_fits, coded_type == type)) +
    scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
    labs(x = "Productive vocabulary size (proportion of items)",
         y = "Grammar score (proportion of items)",
         title = type) +
    scale_color_solarized(name = "") +
    theme(legend.position = "top")
}
```

In our next analysis, we follow up on the relationship between age and grammatical ability found in the "combines" analysis above. In that analysis, we noted that less vocabulary was needed for older children to be marked as combining words. We investigate this pattern in the full Word Form/Complexity item set by splitting data from each language by age. We plot the same curves as above, but separately for children older and younger than the median. 

```{r gram-age_plot_wordforms}
grammar_age_plot("Word Forms")
```


```{r gram-age_plot_complexity}
grammar_age_plot("Complexity")
```
In essentially every language, for both Word Form and Complexity items, we see a higher curve for older children than younger. This finding is consistent with the idea that older children have less vocabulary per unit grammar (mirroring the negative interaction shown for the "combines item."

```{r gram-grammar_diffs}
grammar_areas <- grammar_fits %>%
  group_by(language, instrument, coded_type, age_group) %>%
  summarise(area = sum(mean_fitted * vocab_step, na.rm = TRUE))

ggplot(grammar_areas, aes(x = age_group, y = area, colour = instrument)) +
  facet_wrap(~coded_type) +
  geom_point() +
  geom_line(aes(group = language)) +
  geom_dl(aes(label = language),
          method = list(first.qp, dl.trans(x = x - 0.2),
                        cex = 0.7, fontfamily = .font)) +
  scale_colour_solarized(guide = FALSE) +
  labs(x = "", y = "Area of grammar score\n across vocabulary size") +
  expand_limits(x = 0)
```

This pattern is further summarised in the plot above, where we show the area under the grammar/lexicon curvefor younger and older children. The upward slope of nearly every line demonstrates the consistency of the age effect, which we discuss further below. In addition, there is a trend for age effects to be larger in Complexity rather than Word Forms, suggesting a more syntactic locus for the effect. 

### Individual items

```{r gram-item_models, eval=FALSE}
item_age_model <- function(item_data) {
  glm(value ~ age + 0, data = item_data, family = "binomial")
}

item_vocab_model <- function(item_data) {
  glm(value ~ production_prop + 0, data = item_data, family = "binomial")
}

item_age_vocab_model <- function(item_data) {
  glm(value ~ production_prop + age + 0, data = item_data, family = "binomial")
}

item_models <- grammar_data_coded %>%
  filter(!is.na(value), coded_type != "word_endings") %>%
  group_by(language, form, coded_type, item_id, definition) %>%
  nest() %>%
  mutate(age_model = map(data, item_age_model),
         vocab_model = map(data, item_vocab_model),
         age_vocab_model = map(data, item_age_vocab_model))

item_effects <- item_models %>%
  gather(model_predictors, model, age_model, vocab_model, age_vocab_model) %>%
  mutate(coefs = map(model, tidy),
         AIC = map_dbl(model, AIC),
         converged = map_lgl(model, ~.x$converged)) %>%
  select(-data, -model) %>%
  unnest()

write_feather(item_effects, "data/grammar/item_effects.feather")

# model_comp <- item_effects %>%
#   filter(converged,
#          model_predictors %in% c("vocab_model", "age_vocab_model")) %>%
#   distinct(language, coded_type, item_id, definition, model_predictors, AIC) %>%
#   spread(model_predictors, AIC) %>%
#   mutate(AIC_diff = age_vocab_model - vocab_model,
#          AIC_gain = AIC_diff < 0)
```

```{r gram-item_age_effects}
item_effects <- read_feather("data/grammar/item_effects.feather")

complexity_cats <- items %>%
  filter(type == "complexity") %>%
  distinct(language, form, item_id, complexity_category)

age_effects <- item_effects %>%
  filter(converged, model_predictors == "age_vocab_model", term == "age") %>%
  left_join(complexity_cats) %>%
  mutate(
    coded_type = if_else(
      !is.na(complexity_category) & nchar(complexity_category) > 0,
      complexity_category, coded_type),
    coded_type = factor(
      coded_type,
      levels = c("combine", "word_forms", "morphology", "syntax", "complexity"),
      labels = c("Combine", "Word Forms", "Complexity (morphological)",
                 "Complexity (syntactic)", "Complexity (unknown)"))
  ) %>%
  ungroup() %>%
  arrange(language, form, estimate) %>%
  mutate(order_item = paste(language, item_id, sep = "_"),
         order_item = factor(order_item, levels = order_item))
```

In our final analysis, we examine the individual items on the Word Form and Complexity sections. Given the heterogeneous nature of the CDI instruments, particularly in the Complexity sections, we attempted a more fine-grained item-analysis by classifying items as capturing either more morphological or more syntactic phenomena. Items for which the difference between the simple and complex sentences is in the inflection of a noun or verb (such as *doggie kiss me* / *doggie kissed me*) were coded as morphological. The remainder of the items were coded as Syntactic, since they involved the use of some sentence-level syntactic construction (such as *doggie table* / *doggie on table*).

```{r gram-item_effects_plot}
ggplot(age_effects, aes(x = order_item, y = estimate, colour = coded_type)) +
  facet_wrap(~language, scales = "free", ncol = 3) +
  geom_pointrange(aes(ymin = estimate - 1.96 * std.error,
                      ymax = estimate + 1.96 * std.error), fatten = 1.5) +
  scale_colour_solarized(name = "", guide = guide_legend(ncol = 3)) +
  scale_x_discrete(name = "", breaks = NULL) +
  theme(legend.position = "bottom")
```

We then fit logistic regression models with linear, quaratic, and cubic predictors (as above) separately for every item. The plot above shows the age effect coefficient of each item. In general, age effect were smaller for Word Form items, then Morphological Complexity items, and largest for Syntactic Complexity items, suggesting that more syntactic phenomena likely have greater age contributions. 

<!-- Building on previous analyses that showed a strong relationship between lexical and grammatical development, we added age into this relationship. Across languages, our measures of syntactic development consistently showed greater age modulation than measures of morphological development. Further distinguishing between items that were more reflective of morphology than syntax, we again found greater age effects for more syntactic items. Thus, this analysis provides evidence for a relationship between syntactic development and age _not_ captured by lexical development. -->

## Discussion

We revisited classic findings on the relationship between grammar and the lexicon, further exploring novel questions regarding the role of age in this relation. Our results provid general support for a lexicalist view, in that, in `r num_langs` languages, variance in vocabulary production strongly aligned with variance in grammar. However, we also estimated additional age-related contributions, specifically contrasting the links to morphological forms vs. syntactic constructions, and for different lexical categories. In general, we find that measures of grammar that are more closely aligned with syntax are modulated by age to a greater extent than those reflecting inflectional morphology.

Our analyses suggest new areas of research regarding possible mechanisms driving children's early lexical development and how those mechanisms might support children's transition from single words to more morphosyntactically complex utterances. One possibility is that these developments are dependent on maturational factors that operate on grammatical development in a domain-specific way, independent of lexical-semantic processes. Another possibility is that age-related effects represent more domain-general learning mechanisms, such as attention or working memory, that provide differential support for sentence-level processes than word-internal ones [@gathercole2014]. Future studies should also explore the extent to which lexical and age-related processes are shaped, either independently or in tandem, by features of the learning environments that children experience [e.g., @weisleder2013;@hoff2017].

<!-- FURTHER DEVELOPMENT OF THE INPUT CONNECTION -->

Questions about the nature of morphosyntactic representations in early language have often seemed deadlocked. But by mapping out developmental change across large samples and multiple languages, our findings here challenge theories across the full range of perspectives to more fully describe the mechanistic factors underlying the interaction of vocabulary, grammar, and development. 


