Our second chapter introduces descriptive facts about the dataset as well as details of data formatting and access using the `wordbankr` package (for the R statistical programming environment).

Statistical Methods. Description of our approach to visualization and analysis, including specifications of the statistical models we use for quantifying effect sizes and comparing across studies..

Approach to cross-linguistic analysis. Since comparisons across languages are a key part of our contribution, we outline two different approaches that we use. We identify the difficulties in comparing across CDI forms (including across different languages and even across different forms in a single language). We provide some empirical analysis of how form length affects reporting.

Every form an island. In this approach (EFAI), we compare scale-free measures of effect size across forms, not attempting to bridge the gaps between forms that have different items and different lengths (as well as being designed for different languages).

Universal lemmas. We describe how, for a subset of languages, we identified and checked conceptual mappings that could be used to compare particular concepts across languages, which we call “universal lemmas” (ULs).


# Methods {#methods}

Stuff about the CDI, data collection, population statistics...

We use a number of shorthand terms throughout to make precise references to aspects of the CDI data. These are defined here.


**Intrument**: A specific parent-report survey or questionnaire.

**Item**: A particular question on an instrument. Includes words (with either produces or produces/understands responses), as well as other questions like questions about morphological or syntactic complexity.

**Administration**: A particular instance of an instrument being given to a child.

**Child**: A unique individual.

**Language**: A particular language or language community for which an instrument has been adapted. Note that this definition of language distinguishes e.g. American and British English.

## Database Architecture

[Example data from the CDI norming sample (Fenson et al., 2007). Each row has a unique child identifier, demographics, and word-by-word checklist data.](figures/itembychild.png)

Why use a database to store vocabulary data? Consider the standard format of raw CDI data. Here is a small slice of the original CDI norming data [@fenson1994;@fenson2007]. Each row is a child, each column gives a variable---either a demographic variable or the result of a particular word being administered to a particular child. Although this format is useful for homogeneous administrations of a single instrument, it cannot accommodate multiple instruments, multiple languages, or datasets with different sources or kinds of demographic information. Consolidating data across different instruments is very difficult in this format, and tracking data on children with multiple longitudinal administrations of a single instrument must also be done in an ad-hoc manner. The move to a database format allows far more flexible and programmatic handling of heterogeneous data structures from different sources.

A relational database such as Wordbank is at its heart a series of tables linked by unique identifiers. There are two primary groups of tables in Wordbank. The *common* tables store data that is shared between CDI instruments, including information about *administrations* (individual instances of a form being filled out for a child), and *items* (words and other questions on a form). Then the *instrument* tables store response data for particular CDI instruments. We currently include all items on CDI instruments, including questions about communication, gesture, morphology, and grammar (though in many of the datasets that we archive these non-vocabulary questions have not been digitized so data on these are sparse at present; see Chapters \@ref(label) and \@ref(label)).

One strength of the Wordbank framework is that it allows the storage of subsidiary information about the words that are included in a particular instrument, so that this information can be used in future analyses. For example, information about grammatical and semantic categories or norms like concreteness and imageability could all be appended to particular words. This functionality is not yet present in Wordbank, however. The difficulty of compiling this kind of information for a particular set of words is compounded by the large number of languages that the database includes. We hope that in future this functionality will allow the gradual accumulation of information about the words included in the database.

\subsubsection{Technical details}

Wordbank is constructed using free, open-source tools. The database is a standard \texttt{MySQL} database, managed using \texttt{Python} and \texttt{Django}. Analysis apps are constructed using the \texttt{Shiny} package for \texttt{R}, an open-source statistical programming language.  The code is hosted in a GitHub repository (\url{github.com/langcog/wordbank}) where interested users can browse, leave comments, and contribute modifications.

All data uploaded to Wordbank are open and freely available for download, both through the site itself and through the GitHub repository. The site includes only de-identified data that cannot be linked to the parents and children who provided it. Because of these features, the Stanford Institutional Review Board has determined that the Wordbank project does not constitute human subjects research.

\subsubsection{Cross-linguistic and cross-instrument architecture}

The general philosophy of the CDI has been summarized as ``adaptation, not translation'' \cite{dale2015}. In other words, CDIs are a useful tool for many languages, but the forms differ between languages---words and even whole sections are added, dropped, and modified to ensure that the form captures the details of the particular language for which it is designed. To date, more than 60 adaptations of the original English CDI have been documented \cite{dale2011}. These forms vary widely, including differences in length and intended age range. Some forms include hundreds of items more than the original 680 words on the English Words \& Sentences form; others are so-called ``short forms'' and include only a hundred or a few hundred carefully selected words. Some are designed to capture development from the emergence of language through ages 3--4 years, while others are focused on very early development (like the English Words \& Gestures form, designed for ages 8--18 months). All of these differences make it problematic to compare scores and score distributions across forms, even using percentile ranks, since some instruments will have more or more difficult items than others.

Wordbank is designed so that it can accommodate data from a wide variety of instruments, both within and across languages. Indeed, at the time of writing, the site includes data from more than 42,000 administrations of the CDI across 14 different languages and 24 different instruments. But because of the difficulties in comparison across instruments, our approach to cross-linguistic and cross-instrument data is to provide standardized analyses within each instrument and language, without assuming equivalence across words, instruments, or populations. Thus, our exploratory visualization tools in general do not allow comparison across languages and we urge users to interpret cross-linguistic and cross-instrument differences with caution.\footnote{The only exception to this policy currently is that we allow users to see responses across instruments for individual words, in the Item Trajectories analysis (e.g., the proportion of children who say the word ``cat'' on both Words and Gestures and Words and Sentences forms).} Developing statistical techniques to facilitate these comparisons is a current focus of our research.


## Data

This section gives a broad overview of the data we have available.

### Data Provenance

Many languages are inherited from CLEX.

Contributors are cited on the [Wordbank contributors page](http://wordbank.stanford.edu/contributors.html).

Importing is a process of standardizing idiosyncratic datasets into a common format.

### Overview of the Data

Wordbank curently has data from `r length(unique(items$language))` language communities.

Many of these are in the original Words & Gestures (infant/toddler) / Words & Sentences (child) format, with around 400 items in WG and 700 in the WS.


```{r}
items %>%
  group_by(language, form) %>%
  summarise(items = n(),
            words = sum(type=="word"),
            categories = length(unique(category))) %>%
  filter(form %in% c("WS","WG")) %>%
  datatable
```

Wordbank also includes some other forms that do not fit into this schema, including short forms and vocabulary questionnaires. Some of these are "short forms" with no internal category structure and fewer items overall.

```{r}
items %>%
  group_by(language, form) %>%
  summarise(items = n(),
            words = sum(type=="word"),
            categories = length(unique(category))) %>%
  filter(!(form %in% c("WS","WG"))) %>%
  datatable
```

The number of administrations is variable across instruments and languages, however.

```{r}
admins %>%
  group_by(language, form) %>%
  summarise(n = n(),
            n_unique = length(unique(data_id)),
            min_age = min(age),
            max_age = max(age)) %>%
  datatable
```

## Wordbank Architecture

## Statistical Methods

We use medians instead of means.

We use robust GLMs
