# Vocabulary development {#vocab}

All things vocab size that are over age.

## Norms


## Gender

This analysis examines how vocabulary development differs by children's reported gender, replicating and extending the results of:

Eriksson, M., Marschik, P. B., Tulviste, T., Almgren, M., Pérez Pereira, M., Wehberg, S., ... Gallego, C. (2012). Differences between girls and boys in emerging language skills: Evidence from 10 language communities. British Journal of Developmental Psychology 30, 326–343.

```{r setup, include=FALSE}
library(wordbankr)
library(langcog)
library(dplyr)
library(ggplot2)
library(directlabels)
```

Get administration data and filter to administrations of Words & Sentences that have sex/gender coded.
```{r vocab_admins}
vocab_admins <- get_administration_data() %>%
  select(data_id, language, form, age, sex, production) %>%
  filter(form == "WS", !is.na(sex))
```

Get item information to find the number of items on each language's form.
```{r num_words}
num_words <- get_item_data() %>%
  filter(form == "WS", type == "word") %>%
  group_by(language) %>%
  summarise(n = n())
```

Normalize productive vocabulary size as a proportion of items and calculate median vocabulary size for each language, sex/gender, and age.
```{r vocab_data}
vocab_data <- vocab_admins %>%
  left_join(num_words) %>%
  mutate(production = as.numeric(production) / n) %>%
  group_by(language, sex, age) %>%
  summarise(median = median(production))
```

Plot vocabulary size over age by gender.
```{r plot, fig.width=13, fig.height=8}
min_age <- min(vocab_data$age)
max_age <- max(vocab_data$age)
ggplot(filter(vocab_data),
       aes(x = age, y = median, colour = sex, label = sex)) +
  facet_wrap(~language) +
  geom_line(size = 1) +
  scale_colour_solarized(guide = FALSE) +
  scale_x_continuous(breaks = seq(min_age, max_age, 4),
                     limits = c(min_age, max_age + 5),
                     name = "Age (months)") +
  ylab("Median Productive Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) +
  geom_dl(method = list(dl.trans(x = x + 0.1), "last.points",
                        fontfamily = font, cex = 0.8))
```

## SES

## Birth order
