# Vocabulary Development {#vocab}

This set of analyses examines general patterns of vocabulary development. 

```{r vocab_admins}
num_words <- get_item_data() %>%
  filter(type == "word") %>%
  group_by(language, form) %>%
  summarise(n = n())

vocab_data <- get_administration_data() %>%
  select(data_id, language, form, age, sex, 
         mom_ed, birth_order, production, comprehension) %>% 
  left_join(num_words) %>%
  mutate(no_production = n - production)

min_age <- min(vocab_data$age)
max_age <- max(vocab_data$age)
```


## Norms

Here we examine general patterns of variability across languages. 

### Comprehension

```{r}
source("predictQR_fixed.R")

taus <-  c(0.1, 0.25, 0.5, 0.75, 0.9)
ages <- 8:18

fit_gcrq <- function(x) {
  mod <- try(gcrq(formula = mean ~ ps(age, monotone = 1, 
                                      lambda = 1000), 
                  tau = taus, data = x))
  
  if(inherits(mod, "try-error"))
  {
    return(NA)
  }
  
  return(mod)
}

pred_gcrq <- function(x, mods) {
  mod <- mods[[x$language[1]]]
  
  if (is.na(mod[1])) {
    return(expand.grid(age = x$age, 
                       language = x$language[1], 
                       percentile = as.character(taus*100), 
                       pred = NA))
  } else {
    preds <- predictQR_fixed(mod, 
                             newdata = x) %>%
      data.frame %>%
      mutate(age = x$age, 
             language = x$language) %>%
      gather(percentile, pred, starts_with("X")) %>%
      mutate(percentile = as.character(as.numeric(str_replace(percentile, 
                                                              "X", "")) 
                                       * 100))
    return(preds)
  }
}

comp_data <- vocab_data %>%
  filter(form == "WG") %>%
  mutate(mean = comprehension / n)

comp_models <- comp_data %>%
  split(.$language) %>%
  map(fit_gcrq) 

comp_preds <- comp_data %>%
  group_by(language, age) %>%
  summarise(n = n()) %>%
  data.frame() %>%
  split(.$language) %>%
  map_df(function(x) pred_gcrq(x, comp_models)) 

ggplot(comp_data,
       aes(x = age, y = mean)) +
  facet_wrap(~language) +
  geom_jitter(width = .4, size = 1, alpha = .1) +
  geom_line(data = comp_preds, 
            aes(y = pred, col = percentile, group = percentile)) + 
  scale_colour_solarized(name="Percentile") +
  scale_x_continuous(breaks = seq(8, 18, 4),
                     limits = c(8, 18),
                     name = "Age (months)") +
  ylab("Comprehension Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) + 
  theme(legend.position = "bottom")
```

### Production

Words and Gestures.

```{r}
prod_data <- vocab_data %>%
  filter(form == "WG") %>%
  mutate(mean = production / n)

prod_models <- prod_data %>%
  split(.$language) %>%
  map(fit_gcrq) 

prod_preds <- prod_data %>%
  group_by(language, age) %>%
  summarise(n = n()) %>%
  data.frame() %>%
  split(.$language) %>%
  map_df(function(x) pred_gcrq(x, prod_models)) 

ggplot(prod_data,
       aes(x = age, y = mean)) +
  facet_wrap(~language) +
  geom_jitter(width = .4, size = 1, alpha = .1) +
  geom_line(data = prod_preds, 
            aes(y = pred, col = percentile, group = percentile)) + 
  scale_colour_solarized(name="Percentile") +
  scale_x_continuous(breaks = seq(8, 36, 4),
                     limits = c(8, 36),
                     name = "Age (months)") +
  ylab("Production Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) + 
  theme(legend.position = "bottom")
```

Words and Sentences

```{r}
prod_data <- vocab_data %>%
  filter(form == "WS") %>%
  mutate(mean = production / n)

prod_models <- prod_data %>%
  split(.$language) %>%
  map(fit_gcrq) 

prod_preds <- prod_data %>%
  group_by(language, age) %>%
  summarise(n = n()) %>%
  data.frame() %>%
  split(.$language) %>%
  map_df(function(x) pred_gcrq(x, prod_models)) 

ggplot(prod_data,
       aes(x = age, y = mean)) +
  facet_wrap(~language) +
  geom_jitter(width = .4, size = 1, alpha = .05) +
  geom_line(data = prod_preds, 
            aes(y = pred, col = percentile, group = percentile)) + 
  scale_colour_solarized(name="Percentile") +
  scale_x_continuous(breaks = seq(16, 36, 4),
                     limits = c(16, 36),
                     name = "Age (months)") +
  ylab("Production Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) + 
  theme(legend.position = "bottom")
```


### Comparions of Variability on Comprehension and Production


## Demographic Variables

We examine the patterns of vocabulary growth for individual demographic variables. 

To make these analyses work, we assume that all demographics interact with age in a simple way in the models, such that there is only one coefficient associated with the effect:

age * demographic - demographic

All of these analyses are carried out on production data. 


### Gender

This analysis examines how vocabulary development differs by children's reported gender, replicating and extending the results of:

Eriksson, M., Marschik, P. B., Tulviste, T., Almgren, M., Pérez Pereira, M., Wehberg, S., ... Gallego, C. (2012). Differences between girls and boys in emerging language skills: Evidence from 10 language communities. British Journal of Developmental Psychology 30, 326–343.

An earlier version of this analysis was reported in @frank2016.

Plot vocabulary size over age by gender.

```{r plot, fig.width=13, fig.height=8}
sex_data <- vocab_data %>%
  filter(!is.na(sex), form == "WS") 

sex_means <- sex_data %>%
  group_by(language, sex, age) %>%
  summarise(median = median(production/n))

fit_sex <- function(x) {
  robustbase::glmrob(cbind(production, no_production) ~ age * sex - sex,
                     family = "binomial",
                     data = x)
}

pred_sex <- function(x) {
  x$pred <- predict(sex_models[[x$language[1]]], 
                    newdata = x, type = "response")
  return(x)
}

sex_models <- sex_data %>%
  split(.$language) %>%
  map(fit_sex) 

sex_preds <- expand.grid(language = unique(sex_means$language), 
                         age = min_age:max_age, 
                         sex = unique(sex_means$sex)) %>%
  split(.$language) %>%
  map_df(pred_sex)

ggplot(sex_means,
       aes(x = age, y = median, colour = sex, label = sex)) +
  facet_wrap(~language) +
  geom_point(alpha=.2) +
  geom_line(data = sex_preds, aes(y = pred)) + 
  scale_colour_solarized(name="Sex") +
  scale_x_continuous(breaks = seq(min(sex_means$age), 
                                  max(sex_means$age), 4),
                     limits = c(min(sex_means$age), max(sex_means$age)),
                     name = "Age (months)") +
  ylab("Median Productive Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) + 
  theme(legend.position = "bottom")
```

### Maternal education

We use maternal education as a proxy for socio-economic status. 

```{r mom_ed}
ses_data <- vocab_data %>%
  filter(!is.na(mom_ed), form == "WS") %>%
  mutate(mom_ed = fct_collapse(mom_ed, 
                               `Below Secondary` = c("None","Primary",
                                                     "Some Secondary"),
                               `Secondary` = c("Secondary", "Some College"),
                               `College and Above` = c("College", 
                                                       "Some Graduate", 
                                                       "Graduate")))

ses_means <- ses_data %>%
  group_by(language, mom_ed, age) %>%
  summarise(median = median(production/n))

fit_ses <- function(x) {
  robustbase::glmrob(cbind(production, no_production) ~ 
                       age * mom_ed  - mom_ed,
                     family = "binomial",
                     data = x)
}

pred_ses <- function(x) {
  x$pred <- predict(ses_models[[x$language[1]]], 
                    newdata = x, type = "response")
  return(x)
}

ses_models <- ses_data %>%
  split(.$language) %>%
  map(fit_ses) 

ses_preds <- expand.grid(language = unique(ses_means$language), 
                         age = min_age:max_age, 
                         mom_ed = unique(ses_means$mom_ed)) %>%
  split(.$language) %>%
  map_df(pred_ses)

ggplot(ses_means,
       aes(x = age, y = median, colour = mom_ed, label = mom_ed)) +
  facet_wrap(~language) +
  geom_point(alpha=.2) +
  geom_line(data = ses_preds, aes(y = pred)) + 
  scale_colour_solarized(name="Maternal Education") +
  scale_x_continuous(breaks = seq(min(ses_means$age), 
                                  max(ses_means$age), 4),
                     limits = c(min(ses_means$age), max(ses_means$age)),
                     name = "Age (months)") +
  ylab("Median Productive Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) + 
  theme(legend.position = "bottom")
```


### Birth order

Only `r length(unique(filter(vocab_data,!is.na(birth_order))$language))` languages have birth order data, and only `r length(unique(filter(vocab_data,!is.na(birth_order), form == "WS")$language))` languages have it for WS.

```{r birth_order}
ord_data <- vocab_data %>%
  filter(!is.na(birth_order), form == "WS") %>%
  mutate(birth_order = fct_collapse(birth_order, 
                                    "Third+" = c("Third","Fourth",
                                                 "Fifth","Sixth",
                                                 "Seventh","Eighth")))

ord_means <- ord_data %>%
  group_by(language, birth_order, age) %>%
  summarise(median = median(production/n))

fit_ord <- function(x) {
  robustbase::glmrob(cbind(production, no_production) ~ 
                       age * birth_order  - birth_order,
                     family = "binomial",
                     data = x)
}

pred_ord <- function(x) {
  x$pred <- predict(ord_models[[x$language[1]]], 
                    newdata = x, type = "response")
  return(x)
}

ord_models <- ord_data %>%
  split(.$language) %>%
  map(fit_ord) 

ord_preds <- expand.grid(language = unique(ord_means$language), 
                         age = min_age:max_age, 
                         birth_order = unique(ord_means$birth_order)) %>%
  split(.$language) %>%
  map_df(pred_ord)


ggplot(ord_means,
       aes(x = age, y = median, colour = birth_order, 
           label = birth_order)) +
  facet_wrap( ~language) +
  geom_point(alpha=.2) +
  geom_line(data = ord_preds, aes(y = pred)) + 
  scale_colour_solarized(name="Birth Order") +
  scale_x_continuous(breaks = seq(min(ord_means$age), 
                                  max(ord_means$age), 4),
                     limits = c(min(ord_means$age), max(ord_means$age)),
                     name = "Age (months)") +
  ylab("Median Productive Vocabulary (proportion of total words)") +
  ylim(c(0, 1)) + 
  theme(legend.position = "bottom")
```

### Demographic effect sizes 

How can we compare the size of these different demographic effects? One simple way to do so is to integrate the area between curves. 

The units for this analysis are arbitrary and (right now) can't be compared across languages, since the curves have different lengths. To overcome this limitation, we'd need to look at 

```{r}
momed_levels <- levels(ses_means$mom_ed)

ses_abc <- ses_preds %>%
  spread(mom_ed, pred) %>%
  group_by(language) %>%
  summarise(mom_secondary_ed = sum(`Secondary` - 
                                     `Below Secondary`, na.rm=TRUE), 
            mom_college_ed = sum(`College and Above` - 
                                   `Secondary`, na.rm=TRUE))

sex_abc <- sex_preds %>%
  spread(sex, pred) %>%
  group_by(language) %>%
  summarise(female = sum(Female - Male, na.rm=TRUE))

ord_abc <- ord_preds %>%
  spread(birth_order, pred) %>%
  group_by(language) %>%
  summarise(firstborn = sum(`First` - `Second`, na.rm=TRUE), 
            secondborn = sum(`Second` - `Third+`, na.rm=TRUE))

abcs <- left_join(sex_abc, ord_abc) %>%
  left_join(ses_abc) %>%
  gather(demographic, area_diff, female, firstborn, secondborn, 
         mom_secondary_ed, mom_college_ed) %>%
  mutate(demographic = fct_relevel(demographic, 
                                   c("female", "firstborn", 
                                     "secondborn", "secondary", 
                                     "college")))
```

First look at gender advantage. 

```{r}
ggplot(filter(abcs, demographic == "female"), 
       aes(x = fct_reorder(language, area_diff), y = area_diff)) + 
  geom_point() + 
  coord_flip() + 
  ylab("Advantage for demographic characteristic") + 
  xlab("") + 
  ylim(c(0, 2.5)) + 
  theme(legend.position = "bottom")
```

Visualizing demographic advantages against one another. 

```{r}
n_missing <- abcs %>%
  group_by(language) %>%
  summarise(n_missing = sum(is.na(area_diff)))

ggplot(left_join(abcs, n_missing) %>% filter(n_missing < 3), 
       aes(x = fct_reorder(language, area_diff), 
           group = demographic, col = demographic, 
           y = area_diff)) + 
  geom_line() + 
  ylab("Advantage for demographic characteristic") + 
  xlab("") + 
  ylim(c(0, 4)) + 
  theme(legend.position = "bottom")
```
