# Vocabulary Development {#vocab}

> Far from simply reflecting noise in our measuring instruments or variability in low-level aspects of physiological maturation, the variations that we will document (in vocabulary development) are *substantial,* *stable,* and have their own *developmental course.* Because this variation is substantial, it is critical for defining the boundary between normal and abnormal development; because it is stable, it provides a window onto the correlates and (by inference) the causes of developmental change; and because it has its own developmental course, it can be used to pinpoint critical developmental transitions that form the basis for theories of learning and change. (Bates et al., 1995)

This chapter focuses on one canonical view of CDI data, in which each child is represented by a single vocabulary score: the proportion of words that child knows out of the total in the form. We begin by quantifying the median pattern of vocabulary growth observed in our data; we then turn to characterizing variability across individuals in these data.  

Variability is only meaningful in the case that it is stable; that is, that it reflects signal rather than noise. @bornstein2012 conducted a study in which they collected multiple language measures at 20 and 48 months in a sample of nearly 200 children and used a structural equation model to estimate the stability of a single latent construct, language ability. Overall, essentially all measures related strongly to this latent variable and the coefficient on stability was $r = .84$, suggesting that early language is relatively stable, at least when measured appropriately. Notably, the ELI, a precursor to the CDI, was included in the measures at 20 months and was found to correlate with the 20-month latent construct at $r = .87$.

```{r 3_vocab_admins}
num_words <- items %>%
  filter(type == "word") %>%
  group_by(language, form) %>%
  summarise(n = n())

vocab_data <- admins %>%
  dplyr::select(data_id, language, form, age, sex, 
         mom_ed, birth_order, production, comprehension) %>% 
  left_join(num_words) %>%
  mutate(no_production = n - production)

min_age <- min(vocab_data$age)
max_age <- max(vocab_data$age)

taus <-  c(0.1, 0.25, 0.5, 0.75, 0.9)
ages <- 8:18
```
## Central Tendencies 

One first question we can ask about CDI data is about its central tendency -- the median pattern of vocabulary growth. 

<!-- Linearity: -->
<!-- - left hand side of curves is question of "spurtiness"  - see 7b -->
<!-- - right hand side is question of linearity / exponential growht (number correction vs. ceiling) see appendix E -->


### Commonalities across languages

```{r 3_medians_production_wg, fig.cap="Median production using the Words and Sentences form. Included are only lanugages where there are more than 200 administrations total."}
lang_ns <- admins %>% 
  group_by(language, form) %>% 
  summarise(n_admins = n())

medians <- vocab_data %>%
  group_by(form, language, age) %>%
  summarise(production = median(production), 
            comprehension = median(comprehension), 
            production_prop = median(production/n), 
            comprehension_prop = median(comprehension/n), 
            n = n()) %>%
  left_join(lang_ns) 

ggplot(filter(medians, form == "WG", n_admins > 200), 
       aes(x = age, y = production, col = language)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(se=FALSE, span = 1) +  
  theme(legend.position = "bottom") + 
  xlab("Age (months)") + 
  ylab("Median Production") 
```

Plotted above are the median patterns of growth for early production. Rather than showing proportions, as we will do more standardly throughout the book, here individual item totals are plotted.^[As @eriksson2012 write, "Using raw data assumes that each form is exhaustive, while using percentages assumes that each form is equally exhaustive. Neither is correct and the truth lies somewhere in between."] In general, the median child before the first birthday is reported to produce only a small number of words. These data raise a number of questions about the reliability of early parent reports, which we take up below. 
<!-- This pattern quickly accelerates, however, and (especially for forms that are used beyond 18 months) we observe a clear upward trajectory.  -->
Overall, however, these data accord relatively well with our intuitive sense of early vocabulary development: they reveal that children before their first birthday tend to speak at most only a few words, but that production accelerates across the second year.

```{r 3_medians_production_ws}
ggplot(filter(medians, form == "WS", n_admins > 200), 
       aes(x = age, y = production, col = language)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(se=FALSE, span = 1) +  
  theme(legend.position = "bottom") + 
  xlab("Age (months)") + 
  ylab("Median Production") +
  geom_hline(yintercept = 50, lty = 2)
```

This acceleration is even clearer when looking at production reports from older children using Words and Sentences. In every language, we see the median child is reported to produce 50 words between 16 and 20 months. As we will see below, this analysis masks the tremendous variability apparent during this period. In addition, as we discuss below, languages vary considerably in the absolute number of words reported. Nevertheless, there are still substantial consistencies between languages. 

During the period 24-30 months, we see children beginning to produce a large enough sample of words that curves are leveling out. Presumably this leveling does not reflect a slowing in the rate of acquisition, which many authors assume continues unabated for many years [@bloom2001]. Instead, it reflects the limitations of the CDI instrument, in that there are many possible "more advanced" words that they could be learning, of which only a small subset are represented on any form. See Appendix \@ref{appendix-vocab-estimation) for estimates of how many words children actually might know. 

```{r 3_medians_comprehension_wg}
ggplot(filter(medians, form == "WG", n_admins > 200), 
       aes(x = age, y = comprehension, col = language)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(se=FALSE, span = 1) +  
  theme(legend.position = "bottom") + 
  xlab("Age (months)") + 
  ylab("Median Comprehension") 

mins <- medians %>% 
  filter(n_admins > 200) %>%
  filter(form == "WG", age == 8) 
```

We next turn to comprehension. Comprehension is only queried on the Words and Gestures form. Reported comprehension increases much faster than production; so much so that most parents are reporting that their children understand most words on the form by 18 months. See Chapter \@ref(style) for discussion of the differences in the balance between comprehension and production between children. As with the production data, we see substantial differences across languages in reported vocabulary (see below). 

One striking aspect of comprehension data is how early comprehension is reported.  For example, from 8 months, we see parents reporting medians of `r min(mins$comprehension)` (`r mins$language[mins$comprehension == min(mins$comprehension)]`) and `r max(mins$comprehension)` (`r mins$language[mins$comprehension == max(mins$comprehension)]`) words. Intuitively, to many reseachers (and parents) these high numbers feel unlikely. A spate of recent infancy experiments suggest that in fact, children in the second half of the first year do have some fragmentary representations of many common words available [e.g., @tincoff, @bergelson2012, @bergelson2017]. The representations revealed in these tasks are quite weak -- often amounting to a 2-5% difference in looking to a target on hearing a word uttered -- but, depending on the criterion used by parents, may be what is detected in these early reports. Thus, these estimates may not be as far off as we initially suppose.^[An alternative possibility is that *both* accounts are true, but unconnected: 8-month-olds could in fact know some common words, but parents could be overstimating their vocabulary based on observed behaviors -- in essence, parents could be right, but for the wrong reasons.]

### Cross-language differences

What causes the differences between languages observed above? In this discussion, we examine production on the Words and Sentences form, as the data are the densest and most reliable for this instrument. 

For example, the median pattern of acquisition is slowest in Danish and fastest in Slovak. We consider a range of explanations for this pattern. 

1. Differences could be due to differences in form construction. For example, the Slovak form could have more words, leading to more possibilities for parents to find a word their child produces. Or the Danish form could contain harder words, leading to fewer words being checked. The plot below shows that, while more words on a form clearly is related to a greater number of  *absolute* words checked, the trend is almost entirely absent when examining proportions. 

```{r}
form_size <- left_join(medians, num_words %>% rename(n_words = n)) %>%
  filter(form == "WS", n_admins > 200, age %in% c(24,30)) %>%
  dplyr::select(-comprehension, -comprehension_prop) %>%
  gather(measure, value, production, production_prop)

ggplot(form_size, 
       aes(x = n_words, y = value, col = language))+ 
  geom_point() + 
  facet_grid(measure~age, scales= "free_y") + 
  # geom_text_repel(aes(label = language)) + 
  geom_smooth(method = "lm", aes(group = 1) ) + 
  # scale_color_discrete(guide = FALSE) + 
  xlab("Number of Words on Form") + 
  ylab("Median Production") +
  theme(legend.position = "bottom")
```

In light of this analysis, we can replot the data in terms of proportions. On this view of the data, however, we still see stratification between languages, with Slovak still at the top and Danish at the bottom. What are other possible explanations for this difference?

```{r 3_medians_production_ws_prop}
highlight_langs <- c("Latvian", "English (American)", "Danish")

highlight_data <- filter(medians, form == "WS", n_admins > 200) %>%
         mutate(highlight = ifelse(language %in% highlight_langs, 1, .2))

ggplot(highlight_data, 
       aes(x = age, y = production_prop, col = language)) + 
  geom_point(aes(size = n, alpha = highlight)) + 
  geom_smooth(se=FALSE, span = 1, data = filter(highlight_data, language %in% highlight_langs)) +
  theme(legend.position = "bottom") + 
  scale_alpha_continuous(guide=FALSE) + 
  xlab("Age (months)") + 
  ylab("Median Production Proportion")
  # geom_hline(yintercept = 50, lty = 2)
```

2. Differences could be due to demographic differences across samples. We can examine sample composition in Chapter \@ref(methods) and see that -- to the extent we have access to demographic data -- sample composition does vary in features that affect vocabulary (e.g., maternal education, birth order; see Chapter \@ref(demographics)). 

We are not yet in a position to conduct a full analysis of these differences, controlling for demographics, as demographic differences *also* vary across cultures. But we can examine the difference between Danish, Latvian, and English (American) for example, and note that these differences look quite similar (though noisier) in the female, first-born children of college-educated mothers. Thus, we do not believe that demographic differences fully explain the cross-linguistic differences observed. 

There is some possibility also for cohort effects, in which older sets of data show differences from newer datasets. Most of our data date to the period 2005-2015, but some of the English and Spanish data are older. 

Other differences may relate not to demographics of the sample but to the administration, and these we cannot control for statistically. For example, instructions at administration -- whether written on the form or given by experimenters -- might have been more liberal in the case of Slovak or English (American) samples. Such instructions could have emphasized completeness in reporting vocabulary. Or the circumstances of administration could have been different -- for example, Danish data were collected online while most English data were collected using paper and pencil forms. English (American) data are contributed by many different labs, so there are likely many different administration styles represented.

It's not clear that there is an obvious trend here, as the Danish data for example are relatively recent (as are the Slovak data) and both were collected online using standardized instructions. 

```{r}
select_medians <- vocab_data %>%
  filter(form %in% "WS", language %in% highlight_langs, 
         mom_ed == "College", sex == "Female", birth_order == "First") %>%
  group_by(language, age) %>%
  summarise(production_prop = median(production/n), 
            comprehension_prop = median(comprehension/n), 
            n = n()) 

ggplot(select_medians, 
       aes(x = age, y = production_prop, col = language)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(se=FALSE, span = 1) +
  theme(legend.position = "bottom") + 
  scale_alpha_continuous(guide=FALSE) + 
  xlab("Age (months)") + 
  ylab("Median Production Proportion")
```

2. Differences could be due to cultural or experimental differences in reporting bias. Slovak parents might have a lower criterion for reporting knowledge of a word. Recalling our discussion of these issues in Chapter \@ref(introduction), their model of children's overall competence might be shifted up. This explanation is as an extension of the discussion above of administration and instructions -- perhaps cultural expectations for what it means to be producing a word or cultural expectations for how verbal children are expected to be. 

From this perspective, the disconnect between Mandarin and Cantonese is especially puzzling. These two samples were collected by an overlapping set of coauthors, in similar cultures. 

```{r}
highlight_langs <- c("Mandarin", "Cantonese")

highlight_data <- filter(medians, form == "WS", n_admins > 200) %>%
         mutate(highlight = ifelse(language %in% highlight_langs, 1, .2))

ggplot(highlight_data, 
       aes(x = age, y = production_prop, col = language)) + 
  geom_point(aes(size = n, alpha = highlight)) + 
  geom_smooth(se=FALSE, span = 1, data = filter(highlight_data, language %in% highlight_langs)) +
  theme(legend.position = "bottom") + 
  scale_alpha_continuous(guide=FALSE) + 
  xlab("Age (months)") + 
  ylab("Median Production Proportion")
  # geom_hline(yintercept = 50, lty = 2)
```

4. Finally, differences could be due to true differences in language acquisition. Many researchers working on Danish believe that, due to its phonological properties (MORE HERE), it is truly a difficult language to learn (@bleses).  

In summary, differences between languages in the sheer number of words reported are unlikely to be accounted for by differences in form size or demographic differences between samples. In our view they likely result from a combination of cultural attitudes towards children's language, differences in administration instructions, and real differences in learning across languages. Partialling out these differences would likely require controlled experiments. For this reason, we attempt to avoid interpreting overall differences in vocabulary size wherever possible and limit ourselves to quantities that can be effetively normalized. 


<!-- Do we see a vocabulary explosion in these data? @brent2001 -->


<!-- ### The shape of development -->

<!-- What is the shape of this developmental process? Naively, it appears to be a logistic function, with the characteristic "S" shape describing a rise from floor, a steep acceleration, and a leveling off at ceiling.  -->

<!-- Individual words -->


<!-- @hidaka -->

<!-- @mollica2017 -->



<!-- ### The reliability of early parent reports -->

<!-- Are these first reported instances of production and comprehension reliable? @schneider2016 -->

<!-- ```{r 3_production_wg_early} -->
<!-- ggplot(filter(medians, form == "WG", n_admins > 200),  -->
<!--        aes(x = age, y = production, col = language)) +  -->
<!--   geom_jitter(aes(size = n), width = .2) +  -->
<!--   # geom_smooth(se=FALSE, span = 1) +   -->
<!--   theme(legend.position = "bottom") +  -->
<!--   xlab("Age (months)") +  -->
<!--   ylab("Median Production") +  -->
<!--   ylim(0,30) + -->
<!--   xlim(7,14) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- ggplot(filter(admins, form == "WG", age == 8) %>% -->
<!--          left_join(lang_ns) %>% -->
<!--          filter(n_admins > 200),  -->
<!--        aes(x = production)) +  -->
<!--   # geom_point(aes(size = n)) +  -->
<!--   geom_histogram(binwidth = 2) +  -->
<!--   facet_wrap(~language) + -->
<!--   xlab("Production at 8 months") +  -->
<!--   xlim(0,50) +  -->
<!--   ylim(0,20) -->
<!-- ``` -->

## Variability between individuals

One of the most important features of early vocabulary development is its variability [@fenson1994]. In this section, we examine this variability. To do so, we switch to a view of the data that reveals the full range of variation across the samples in the database. Examining Words and Sentences data, we can see that across every language in the database there is a tremendous range of vocabulary sizes reported. How systematic is this variability?

<!-- ```{r 3_production_wg} -->
<!-- wg_prod_data <- vocab_data %>% -->
<!--   filter(form == "WG") %>% -->
<!--   mutate(mean = production / n) -->

<!-- wg_prod_models <- wg_prod_data %>% -->
<!--   split(.$language) %>% -->
<!--   map(fit_gcrq)  -->

<!-- wg_prod_preds <- wg_prod_data %>% -->
<!--   group_by(language, age) %>% -->
<!--   summarise(n = n()) %>% -->
<!--   data.frame() %>% -->
<!--   split(.$language) %>% -->
<!--   map_df(function(x) pred_gcrq(x, wg_prod_models))  -->

<!-- ggplot(wg_prod_data, -->
<!--        aes(x = age, y = mean)) + -->
<!--   facet_wrap(~language) + -->
<!--   geom_jitter(width = .4, size = 1, alpha = .1) + -->
<!--   geom_line(data = wg_prod_preds,  -->
<!--             aes(y = pred, col = percentile, group = percentile)) +  -->
<!--   scale_colour_solarized(name="Percentile") + -->
<!--   scale_x_continuous(name = "Age (months)",  -->
<!--                      breaks = seq(8, 20, 4), -->
<!--                      limits = c(8, 20)) + -->
<!--   scale_y_continuous(breaks = c(0, .5, 1), lim = c(0, 1),  -->
<!--                      name = "Production Vocabulary") + -->
<!--   theme(legend.position = "bottom") -->
<!-- ``` -->

```{r 3_production_ws}
ws_prod_data <- vocab_data %>%
  filter(form == "WS") %>%
  mutate(mean = production / n)

ws_prod_models <- ws_prod_data %>%
  split(.$language) %>%
  map(fit_gcrq) 

ws_prod_preds <- ws_prod_data %>%
  group_by(language, age) %>%
  summarise(n = n()) %>%
  data.frame() %>%
  split(.$language) %>%
  map_df(function(x) pred_gcrq(x, ws_prod_models)) 

ggplot(ws_prod_data,
       aes(x = age, y = mean)) +
  facet_wrap(~language) +
  geom_jitter(width = .4, size = 1, alpha = .02) +
  geom_line(data = ws_prod_preds, 
            aes(y = pred, col = percentile, group = percentile)) + 
  scale_colour_solarized(name="Percentile") +
  scale_x_continuous(name = "Age (months)", 
                     breaks = seq(16, 36, 4),
                     limits = c(16, 36)) +
  scale_y_continuous(breaks = c(0, .5, 1), lim = c(0, 1), 
                     name = "Production Vocabulary") +
  theme(legend.position = "bottom")
```

<!-- ```{r 3_comprehension} -->
<!-- comp_data <- vocab_data %>% -->
<!--   filter(form == "WG") %>% -->
<!--   mutate(mean = comprehension / n) -->

<!-- comp_models <- comp_data %>% -->
<!--   split(.$language) %>% -->
<!--   map(fit_gcrq)  -->

<!-- comp_preds <- comp_data %>% -->
<!--   group_by(language, age) %>% -->
<!--   summarise(n = n()) %>% -->
<!--   data.frame() %>% -->
<!--   split(.$language) %>% -->
<!--   map_df(function(x) pred_gcrq(x, comp_models))  -->

<!-- # breaks = seq(8, 18, 4), -->
<!-- #                      limits = c(8, 18), -->
<!-- ggplot(comp_data,  -->
<!--        aes(x = age, y = mean)) + -->
<!--   facet_wrap(~language) + -->
<!--   geom_jitter(width = .4, size = 1, alpha = .1) + -->
<!--   geom_line(data = comp_preds,  -->
<!--             aes(y = pred, col = percentile, group = percentile)) +  -->
<!--   scale_colour_solarized(name="Percentile") + -->
<!--   scale_x_continuous(name = "Age (months)",  -->
<!--                      seq(8, 18, 4), -->
<!--                      limits = c(8, 18)) + -->
<!--   scale_y_continuous(breaks = c(0, .5, 1), lim = c(0, 1),  -->
<!--                      name = "Comprehension Vocabulary") + -->
<!--   theme(legend.position = "bottom") -->
<!-- ``` -->




<!-- Stiched across forms. See \@ref(appendix-stitching). -->

<!-- ```{r 3_production_stitched} -->

<!-- ``` -->




<!-- ## Quantifying Variability -->


```{r 3_eng_prod}
eng_prod_data <- vocab_data %>%
  filter(form == "WS", language == "English (American)") %>%
  mutate(mean = production)

eng_prod_model <- list()
eng_prod_model$`English (American)` <- eng_prod_data %>%
  fit_gcrq

eng_prod_preds <- eng_prod_data %>%
  group_by(age, language) %>%
  summarise(n = n()) %>%
  data.frame() %>% 
  pred_gcrq(eng_prod_model)

ggplot(eng_prod_data,
       aes(x = age, y = mean)) +
  geom_jitter(width = .4, size = 1, alpha = .1) +
  geom_line(data = eng_prod_preds, 
            aes(y = pred, col = percentile, group = percentile)) + 
  scale_colour_solarized(name="Percentile") +
  scale_x_continuous(breaks = seq(16, 30, 4),
                     limits = c(16, 30),
                     name = "Age (months)") +
  ylab("Production (Number of Words)") +
  ylim(c(0, 680)) + 
  theme(legend.position = "bottom")
```

As an example, we zoom in on the English (American) production data from Words and Sentences form. Consider just a single age group, 24-month-olds. The distribution of vocabularies across children is far from normally distributed, with many children at the very bottom of the scale and almost as many at the top. Quite a few two-year-olds on their second birthday are producing only a handful of words (or at least their parents say they are) and others are producing more than 600 listed on the form, with the total number higher than that even.
<!-- ^[It will quickly get tiresome to acknowledge ceiling effects and parent report biases in every sentence, so we will acknowledge them up front and then mention them only when relevant throughout.]  -->

```{r v_histogram}
ggplot(filter(eng_prod_data, age == 24), 
       aes(x = production)) + 
  geom_histogram(aes(y = ..density..), binwidth=30) + 
  geom_density(col = "red", aes(y = ..density..)) +
  xlim(0,680) + 
  xlab("Number of words produced")
```

One way to describe these data is to consider the relationship of the variance to the central tendency. The "coefficient of variation" (CV) is a common measure used for this purpose: 

$$CV = \frac{\sigma}{\mu}$$
This statistic allows standardized comparison of variability across measurements with different scales, an important concern when we want to compare forms with very different numbers of vocabulary items. For example, for two-year-olds, the mean productive vocabulary is `r round(mean(eng_prod_data$production[eng_prod_data$age == 24]))` words, and the standard deviation is `r round(sd(eng_prod_data$production[eng_prod_data$age == 24]))`, words, leading to a CV of `r signif(sd(eng_prod_data$production[eng_prod_data$age == 24])/mean(eng_prod_data$production[eng_prod_data$age == 24]), digits = 2)`. 

But, as seen in Figure \@ref(fig:v_histogram), the distribution of productive vocabulary scores is far from normal; this trend is even more apparent at the younger and older ages. Thus, a non-parametric approach is more appropriate. Accordingly, we compute the MADM statistic, the non-parametric equivalent of the CV. In MADM, the mean $\mu$ is replaced by the median ($m(x)$, and the standard deviation $\sigma$ is replaced by the mean absolute deviation (which captures how far away values are from the median):

$$MADM(x) = \frac{\frac{1}{n} \sum_{i = 1..n}{|x_i - m(x)|}}{m(x)}$$


```{r 3_madm1}

madm_data <- vocab_data %>%
  filter(form %in% c("WS","WG")) %>%
  group_by(language, form, age) %>%
  summarise(madm_p = madm(production), 
            madm_c = madm(comprehension),
            n = n()) %>%
  filter(n > 20) 

ggplot(filter(madm_data, language == "English (American)"), 
       aes(x = age, y = madm_p, col = form)) + 
  geom_point() + 
  geom_smooth(aes(group = 1)) + 
  ylim(0,2) +
  xlim(8,30) + 
  ylab("MADM (MAD / median)") + 
  xlab("Age (months)") + 
  scale_color_solarized(name = "Form")
```

In American English production, this ratio is actually close to 1 from age one until almost age two, suggesting that the standard *difference* from the median is actually as big as the median itself! The decline begins before variability is substantially truncated by the ceiling of the form, suggesting that variability between kids is really highest before the second birthday. 

Imagine groups of three children. A group where one produced 30 words, one produced 100, and another produced 170 would have a MADM of `r round(madm(c(33,100,180)), digits = 2)`. In contrast, one where they were more closely grouped -- say 70, 100, 130 -- would have a MADM of `r round(madm(c(70,100,130)), digits = 2)`.

```{r 3_madm2}
ggplot(madm_data, aes(x = age, y = madm_p, col = language)) + 
  geom_jitter(aes(size = n), alpha = 1) + 
  geom_smooth(aes(group  = 1)) + 
  ylim(0,2) + 
  ylab("MADM (Production)") + 
  xlab("Age (months)") + 
  scale_color_solarized(name = "") + 
  theme(legend.position = "bottom")
```

The next figure shows MADM across languages and instruments. This similarity in variability structure is quite striking, such that between the first and second birthdays, language is remarkably variable. Yet this variability is quite *consistent*! We summarise the MADM in the second year of life by taking its mean. This mean is close to 1 for almsot every language and form for which we have data. 

One question that could be raised regarding the analysis above is the extent to which variability is caused by *variability across children* vs. *variability in reporting*. The extreme values seen in the English data, for example, could easily be the result of a mixture of lazy parents who stopped answering the form with overly dilligent parents who misunderstood and checked every box for a word they thought the child had been exposed to. Several arguments speak against this account, however. First, to the extent these biases are the source of variability, they are extremely *consistent* across languages -- which, recall, is the exact opposite argument from the one we considered above (where parent dilligence was supposed to be variable enough across samples to lead to differences between languages). In addition, note that the 
In sum, although there are certainly some reporting biases represented in the data, these results do not appear likely to be an artifact of reporting bias.

```{r 3_madm_summary}
madm_summary_prod <- madm_data %>%
  filter(language != "Kigiriama") %>%
  group_by(language, form) %>%
  filter(age >= 12, age <= 24) %>%
  summarise(madm_mean = mean(madm_p), 
            madm_ci_lower = mean(madm_p) - ci.t(madm_p),
            madm_ci_upper = mean(madm_p) + ci.t(madm_p)) 

ggplot(madm_summary_prod, aes(x = language, y = madm_mean, col = form)) + 
  geom_pointrange(aes(ymin = madm_ci_lower, ymax = madm_ci_upper), 
                  position = position_dodge(width = .1)) + 
  ylim(0,1.5) + 
  coord_flip() + 
  ylab("Language") + 
  xlab("MADM (Production 12-24 mo)") +
  geom_hline(aes(yintercept = 0), lty = 2, color = "gray") + 
  # geom_hline(aes(yintercept = 1), lty = 2, color = "gray") + 
  scale_colour_solarized()
```


```{r 3_madm_comprehension}
ggplot(filter(madm_data, form == "WG"), 
       aes(x = age, y = madm_c, col = language)) + 
  geom_jitter(aes(size = n), alpha = 1) + 
  geom_smooth(aes(group  = 1)) + 
  ylim(0,2) + 
  ylab("MADM (Comprehension)") + 
  xlab("Age (months)") + 
  scale_color_solarized(name = "") + 
  theme(legend.position = "bottom")
```

Ae can complete the same analysis using comprehension data. Here we see a gradual decrease in variability throughout development. The intercept for the 12-18 month period appears to be lower than that observed in production, despite (or because of?) the higher scores. One speculative explanation for this difference would be the tremendous differences in speech-motor development (as well as general differences in loquacity) between toddlers. This variability is then reflected in production.

This observation matches one made by @mayor2012, namely that production vocabulary appears more idiosyncratic in distribution of words than comprehension vocabulary. 

```{r 3_madm_summary_comprehension}
madm_summary_comp <- madm_data %>%
  filter(form == "WG", language != "Kigiriama", age < 19, age > 11) %>%
  group_by(language, form) %>%
  summarise(madm_mean = mean(madm_c), 
            madm_ci_lower = mean(madm_c) - ci.t(madm_c),
            madm_ci_upper = mean(madm_c) + ci.t(madm_c)) 

ggplot(madm_summary_comp, aes(x = language, y = madm_mean)) + 
  geom_pointrange(aes(ymin = madm_ci_lower, ymax = madm_ci_upper)) + 
  ylim(0,1.5) + 
  coord_flip() + 
  ylab("Language") + 
  xlab("MADM (Comprehension)") +
  geom_hline(aes(yintercept = 0), lty = 2, color = "gray") 
  # geom_hline(aes(yintercept = 1), lty = 2, color = "gray")
```

## Discussion

In sum, we observed a striking consistency in the individual variability of children's vocabulary during their second year. Across languages and forms, it appears to be the norm that toddlers vary. 

This variability is found with striking consistency across our data. 
<!-- Nevertheless, it is still possible that  -->

@feldman2000 discusses the possibility that some portion of the observed variability in her data as well as the CDI norming data [@fenson1994] is due to reporting issues. In their response, @fenson2000 cite an unpublished dataset comparing lab-based elicited production and parent report for 36 object names and show that the standard deviation is greater than the mean on both measures, suggesting that the variability found in CDI scores is not artifactual. 



```{r 3_save_covs}
prod <- madm_summary_prod %>%
  group_by(form) %>%
  summarise(prod_madm =  sd(madm_mean) / mean(madm_mean))

comp <- madm_summary_comp %>%
  ungroup %>%
  summarise(comp_madm =  sd(madm_mean) / mean(madm_mean))

cvs <- data_frame(measure = c("MADM 12-24mo WG production", 
                       "MADM 12-24mo WS production", 
                       "MADM 8-18mo WG comprehension"), 
           cv = c(prod$prod_madm, comp$comp_madm))

write_feather(cvs, "data/cvs/madm_cvs.feather")
```

