# Measures of Variability {#appendix-variability}

In Chapter \@ref{vocab}, we switch between parametric and non-parametric measures of variability (CV vs. MADM) and Cohen's $d$ vs. MMAD. Here we show that these are very similar in the limit with a large amount of data, although they can produce quite different answers for individual data points, especially those that are at the floor or ceiling of the particular form. 


```{r}
num_words <- items %>%
  filter(type == "word") %>%
  group_by(language, form) %>%
  summarise(n = n())

vocab_data <- admins %>%
  dplyr::select(data_id, language, form, age, sex, 
         mom_ed, birth_order, production, comprehension) %>% 
  left_join(num_words) %>%
  mutate(no_production = n - production)

```


```{r 3_madm_normal}
ratios <- vocab_data %>%
  filter(form %in% c("WS","WG")) %>%
  group_by(language, form, age) %>%
  filter(n() > 20) %>%
  summarise(mmad = mmad(production), 
            madm = madm(production), 
            cv = cv(production), 
            d = d(production), 
            n = n())
```


```{r}
ggplot(ratios, aes(x = madm, y = cv)) + 
  geom_point(aes(size = n, col = language), alpha = .3) + 
  geom_smooth(method = "lm") +
  ylim(0,3) + xlim(0,3) +
  geom_abline(lty = 2) +
  ylab("Cohen's d") + 
  xlab("MADM") + 
  scale_color_discrete(guide = FALSE)
```


```{r}
ggplot(ratios, aes(x = mmad, y = d)) + 
  geom_point(aes(size = n, col = language), alpha = .3) + 
  geom_smooth(method = "lm") +
  ylim(0,3) + xlim(0,3) +
  geom_abline(lty = 2) +
  ylab("Cohen's d") + 
  xlab("MADM") + 
  scale_color_discrete(guide = FALSE)
```

