# Stitching Across Forms {#appendix-stitching}

Because we use different forms for different ages, there are sometimes good reasons to combine data across forms to get a broader range of ages in a particular analysis. We call this combination "stitching." This appendix provides some motivation for the practice. 


## Item Stitching

For arbitrary items, stitching across English WS and WG tends to look pretty good. 

```{r appstitch-example}
target_items <- c("dog","table","run")

eng_wg <- get_instrument_data(language = "English (American)",
                              form = "WG", 
                              iteminfo = TRUE, 
                              items = items %>%
                                filter(language == "English (American)", 
                                       form == "WG", 
                                       uni_lemma %in% target_items)  %>%
                                pull(item_id)) %>%
  filter(uni_lemma %in% target_items)

eng_ws <- get_instrument_data(language = "English (American)",
                              form = "WS", 
                              iteminfo = TRUE, 
                              items = items %>%
                                filter(language == "English (American)", 
                                       form == "WS", 
                                       uni_lemma %in% target_items)  %>%
                                pull(item_id)) %>%
  filter(uni_lemma %in% target_items)

eng_items <- bind_rows(eng_wg, eng_ws) %>%
  left_join(admins) %>%
  group_by(age, form, uni_lemma) %>%
  summarise(produces = mean(value == "produces", na.rm=TRUE))

ggplot(eng_items, 
       aes(x = age, y = produces, col = form)) + 
  facet_wrap(~uni_lemma) +
  geom_point() + 
  geom_smooth()
```


## Form Stitching

We can "stitch" across WS and WG forms to get a fuller picture of production across ages. There are several ways to accomplish this. The simplest is to use proportions for each form. 
This isn't particularly good.

```{r appstitch-eng_prod}
num_words <- items %>%
  filter(type == "word") %>%
  group_by(language, form) %>%
  summarise(n = n())

vocab_data <- admins %>%
  select(data_id, language, form, age, sex, 
         mom_ed, birth_order, production, comprehension) %>% 
  left_join(num_words) %>%
  mutate(no_production = n - production)

eng_prod_data <- vocab_data %>%
  filter(language == "English (American)") %>%
  mutate(mean = production / n)

ggplot(eng_prod_data,
       aes(x = age, y = mean, col = form)) +
  facet_wrap(~language) +
  geom_jitter(width = .4, size = 1, alpha = .1) +
  geom_smooth() + 
  # geom_line(data = ws_prod_preds,
  #           aes(y = pred, col = percentile, group = percentile)) +
  .scale_colour_discrete(name="Percentile") +
  scale_x_continuous(name = "Age (months)",
                     breaks = seq(8, 30, 4),
                     limits = c(8, 30)) +
  scale_y_continuous(breaks = c(0, .5, 1), lim = c(0, 1),
                     name = "Production Vocabulary") +
  theme(legend.position = "bottom")
```

You can also stitch item by item, this is trickier but better. 

## Item-Level Form Stitching
